---
title: "Introduction to R"
output: html_notebook
editor_options: 
  chunk_output_type: inline
---
I recommend the following R tutorials & short articles.

- R Learning Path: From beginner to expert in R in 7 steps https://www.kdnuggets.com/2016/03/datacamp-r-learning-path-7-steps.html
- RStudio Cheat Sheets https://www.rstudio.com/resources/cheatsheets/
- R Notebook Cheat Sheet https://www.rstudio.com/wp-content/uploads/2015/02/rmarkdown-cheatsheet.pdf
- https://www.tutorialspoint.com/r/index.htm
- http://www.statmethods.net/
- http://www.r-tutor.com/
- https://github.com/kabacoff/RiA2
- R for Everyone (you can access this book from through K-State Library http://guides.lib.k-state.edu/az.php?q=safari)


### Variables
```{r}
# declare variables

myString <- "Hello, World!"
myNumber <- 1000
print (myString)
print (myNumber)
```

### Data Types
```{r}
a <- c(1,2,5.3,6,-2,4) # numeric vector
b <- c("one","two","three") # character vector
c <- c(TRUE,TRUE,TRUE,FALSE,TRUE,FALSE) #logical vector

a[c(2,4)] # 2nd and 4th elements of vector
```

### Operators (Arithmetic Operators)

+	addition
-	subtraction
*	multiplication
/	division
^ or **	exponentiation
x %% y	modulus (x mod y) 5%%2 is 1
x %/% y	integer division 5%/%2 is 2

### Operators (Logical Operators)

<	less than
<=	less than or equal to
>	greater than
>=	greater than or equal to
==	exactly equal to
!=	not equal to
!x	Not x
x | y	x OR y
x & y	x AND y
isTRUE(x)	test if X is TRUE


### Decision making (if statements)
```{r}
x <- 10

if(x > 5) {
   print("x is a big number")
} else {
   print("x is smaller than 5")
}
```

```{r}
x <- 5

if(x > 5) {
   print("x is greater than 5")
} else if(x < 5) {
  print("x is smaller than 5")
} else {
   print("x is equal to 5")
}
```

```{r}
x <- c("what","is","truth")

if("Truth" %in% x) {
   print("Truth is found the first time")
} else if ("truth" %in% x) {
   print("truth is found the second time")
} else {
   print("No truth found")
}
```


### Loops

```{r}
x <- c("I", "Love", "You")
for ( i in x) {
   print(i)
}
```

```{r}
for (year in 2010:2015){
  print(paste("The year is", year))
}
```

```{r}
for (i in 1:10) {
  if (i %% 2){
    next
  }
    print(i)
}
```

```{r}
for (i in 1:10) {
  if (!i %% 2){
    next
  }
    print(i)
}
```


```{r}
# Create the data frame.
people <- 	data.frame(
   gender = c("Male", "Male","Female"), 
   height = c(152, 171.5, 165), 
   weight = c(81,93, 78),
   Age = c(42,38,26)
)
print(people)
```

```{r}
# example of a list with 4 components - 
# a string, a numeric vector, a matrix, and a scaler 
w <- list(name="Fred", mynumbers=10, mymatrix=5, age=5.3)

print (w)
```

### Getting and Setting the Working Directory
```{r}
# Get and print current working directory.
print(getwd())

# Set current working directory.
setwd("C:/Users/bchae/Documents/R/665")
# setwd("C:/Users/bchae/Documents/R/670")

# Get and print current working directory.
print(getwd())
```

### Exporting & Importing data
```{r}
data <- read.csv("tips.csv")
print(data)
```

```{r}
write.csv(data, "tips_copy.csv")
```

### View data
```{r}
head(data)
```

```{r}
dim(data)
```

```{r}
str(data)
```

```{r}
mean(data$tip)
```

```{r}
max(data$tip)
```
```{r}
summary(data)
```

### Data transformation: Selecting columns & Changing data types

dplyr and tidyr are two popular R package for data manipulation 
https://www.rstudio.com/wp-content/uploads/2015/02/data-wrangling-cheatsheet.pdf

```{r}
library(dplyr)
```

```{r}
data <- read.csv("titanic_train.csv")
head(data)
```

```{r}
str(data)
```

```{r}
df = select(data, Survived, Pclass, Sex, Age, Fare, Embarked)
head(df)
```


```{r}
str(df)
```

Embarked has empty cells
```{r}
df[df==""] <- NA
str(df)
```

Check NA
```{r}
sapply(df, function(x) sum(is.na(x)))
```

```{r}
df1 <- na.omit(df)
sapply(df1, function(x) sum(is.na(x)))
```

```{r}
dim(df1)
```


```{r}

df1$Sex <- as.numeric(df1$Sex)
#df$Embarked <- as.numeric(df$Embarked)
str(df1)

```


```{r}
#df$Sex <- as.numeric(as.factor(df$Sex))
df1$Embarked <- as.numeric(df1$Embarked)
str(df1)
```

```{r}
head(df1)
```

save the data after cleaning & transforming
```{r}
write.csv(df1, "cleanedata.csv")
```

